**Title:** Lake Photic Zone Temperature across the Conterminous United States

### 2020-10-08 manuscript/manuscript.Rmd

In addition to climate, there are additional factors clearly impacting lake temperature (e.g., surrounding land use, lake depth, size, and configuration, and elevation).   To test for the relative importance of lake morphometry and surrounding landscape, we used the R packages `lakemorpho` to calculate a suite of lake morphometry metric and `elevatr` to access digital elevation models for each lake [@hollister2017lakemorpho; @hollister2017elevatr]. Additionally, the National Land Cover Database (NLCD) was our source for land cover data.  Specifically, we calculated the percent impervious surface of a 3000m buffer for each lake.  Lakes with partials buffers falling outside the US were excluded.

@@

Random forest modelling was used to not only develop a predictive model of photic zone temperature, but also used to aid in variable selection and to calculate relative variable importance.  Random forest is a machine learning method that builds a consensus prediction from the assemblage of multiple tree models (here specifically 10,000 trees for the final model and 1,000 trees for the variable selection models).  Each individual tree model is constructed from a subset of the full data set and a subset of all predictor variables.  All random forest modelling was conducted in R, version 4.0.2 with the randomForest package [@R2020; @liaw2002rf].  Model performance is reported as mean square error and adjusted R^2^.

@@

In addition to the several derived air temperature variables, we included land-use/land cover variables in our initial variable selection process. Specifically, we calculated the percent impervious surface for a 3km lake buffer and a measure of shoreline development. These variables were included based on the hypothesis that higher amounts of development and therefore impervious surface surrounding a lake would lead to higher temperatures in lakes. Yet neither of these variables were selected in the final model. Even though the land-use variables were not selected that does not mean development and impervious are not impactful. This urban-heat effect on lakes may have been adequately captured in the average ambient air temperature. Therefore, making the land-use variables redundant. Regardless these variables did not independently contribute to the model’s accuracy.

>>>>

In addition to climate, there are additional factors clearly impacting lake temperature (e.g., surrounding land use, lake depth, size, and configuration, and elevation).   To test for the relative importance of lake morphometry and surrounding landscape, we used the R packages `lakemorpho` to calculate a suite of lake morphometry metrics and `elevatr` to access digital elevation models for each lake [@hollister2017lakemorpho; @hollister2017elevatr]. Additionally, the National Land Cover Database (NLCD) was our source for land cover data.  Specifically, we calculated the percent impervious surface of a 3000m buffer for each lake.  Lakes with partials buffers falling outside the US were excluded.

@@

Random forest modelling was used to not only develop a predictive model of photic zone temperature, but also used as a means of variable selection and to calculate relative variable importance.  Random forest is a machine learning method that builds a consensus prediction from the assemblage of multiple tree models (here specifically 10,000 trees for the final model and 1,000 trees for the variable selection models).  Each individual tree model is constructed from a subset of the full data set and a subset of all predictor variables.All random forest modelling was conducted in R v XXXX (CITE) with the randomForest package (CITE).  Model performance is reported as mean square error and adjusted R.

@@

# Alternate to the above paragraph

The final model included average ambient air temperature and the average temperature of the prior 30 days,yet the 3-day and 7-day prior to the sampling date averages were not selected. It is likely that the three and seven day averages are not providing the model with unique information because of the focus on photic zone temperature. The upper 2m of the water column is more responsive to short term temperature changes compared to the whole lake, thus the 3 and 7 day temperatures are likely redundant with the sample date temperature. Yet the longer term 30-day average does have an impact on lake photic zone temperature, more specifically, this measures the long term thermal heating happening at a site. The 30-day average provides us with information about the temperature intensity leading up to the sample date. This is a separate measure from how far along a lake is into summer; that information would be captured in the date parameter.

In addition to the several derived air temperature variables, we included land-use/land cover variables in our initial variable selection process. Specifically, we calculated the percent impervious surface for a 3km lake buffer and a measure of shoreline development. These variables were included based on the hypothesis that higher amounts of development and therefore impervious surface surrounding a lake would lead to higher temperatures in lakes. Yet neither of these variables were selected in the final model, although lake shoreline length was selected. Even though the land-use variables were not selected that does not mean development and impervious are not impactful. This urban-heat effect on lakes may have been adequately captured in the average ambient air temperature. Therefore, making the land-use variables redundant. Regardless these variables did not independently contribute to the model’s accuracy. It is important to note that including other variables that impact lake temperature, such as solar radiation, would likely improve the accuracy of the model.  However, these variables are not available for all lakes in the conterminous United States.
